Enabling the Use of GPU Virtualization in Cloud Environments

### 2015-10-21 paper.tex
primera version introduccion

\section{\uppercase{Introduction}}
\label{sec:introduction}

>>>>

\section{\uppercase{Introduction}}
\label{sec:introduction}
The growth in the use of coprocessor and accelerators has been continuous
in the past several years. In the most recent TOP500
ranking~\cite{nov14top500},  dated June 2015, 90 supercomputers
were equipped with accelerators of some sort.  These devices offer
performance improvements over traditional processors for
computational kernels which match their architecture; moreover, 
they are more energy efficient achieving an important MFlop/Watt ratio.

The generalized adoption of coprocessors to accelerate general-purpose 
codes started with the release of the CUDA programming model for
NVIDIA GPUs~\cite{cuda65}. This libraries include both high- and low-level application programming interfaces (APIs)
to allow the use of NVIDIA devices as general-purpose
accelerators.  
%Prior to that, researchers matched 
%GPU architectures to computational kernels, but these had to
%use graphics-oriented APIs such as OpenGL~\cite{opengl} or
%Cg~\cite{cg} to do so. 
Following CUDA, OpenCL~\cite{opencl} appeared as an attempt
to offer a cross-vendor solution. The third-generation programming
model for accelerators, based on compiler directives, started with
OpenACC~\cite{openaccweb}. This standard defines a collection of 
directives to facilitate platform-independent accelerator
usage. This generation is also completed with other task-oriented, directive-based programming model such as OmpSs~\cite{ompssweb} 
which can target directly CPUs or accelerators and recently, OpenMP~\cite{openmp4} incorporated some
accelerator-oriented directives in its version 4.
%, although 
%no available compiler supports them to date.

The current approach for creating clusters with GPUs has been to
populate every compute node with one or more of these devices. 
However, these configurations present a low
utilization rate of the computational resources available in the
hardware accelerators, not only because of mismatches between the application's type of parallelism
and the coprocessor architecture, but also because of applications can be hardly parallelized more than 
80\% of the total code. 
In order to address this problem, remote virtualization was proposed.
Among the virtualization frameworks, the most
prominent is \rcuda~\cite{tonithesis,toniparco}, which
enables cluster configurations with fewer GPUs than nodes.  With this
approach, GPU-equipped nodes act as GPGPU (general-purpose GPU) servers for the rest of the
compute nodes of the cluster, enabling a CUDA-sharing solution that potentially achieves 
 a higher overall GPU load in the system. Compared with 
other CUDA and OpenCL virtualization frameworks (e.g.
DS-CUDA~\cite{dscuda}, vCUDA~\cite{vcuda}, VOCL~\cite{vocl}, and SnuCL~\cite{snucl}),
 \rcuda\ is a mature production-ready framework
 that offers support for the latest CUDA revisions and provides wide
coverage of current GPGPU APIs.
